---
title: "Boxplot Exercises"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The InsectSprays data set measures the counts of insects in agricultural experimental units treated with different insecticides. This dataset is included in R, and you can examine it by typing:  
```{r}
data <- InsectSprays
head(data)
```  
There are two ways of drawing boxplots in R.  
```{r}
boxplot(split(data$count,data$spray))
```
```{r}
boxplot(data$count ~ data$spray)
```  

###Exercise 1
Which spray is the most effective (lowest median)?  
C  

###Exercise 2
Let's consider a random sample of finishers from the New York City Marathon in 2002. This dataset can be found in the UsingR package. Load the library and then load the nym.2002 dataset.  
```{r message=FALSE}
library(UsingR)
library(dplyr)
runners <- nym.2002
```
Use boxplots and histograms to compare the finishing times of males and females. Which of the following best describes the difference?  
```{r}
boxplot(split(runners$time,runners$gender)) 
data <- runners %>% select(time,gender)
split <- split(data,data$gender)
hist(split$Female$time, main="Female")
hist(split$Male$time, main="Male")

```
Male and females have similar right skewed distributions with the former, 20 minutes shifted to the left

